---
title: "Homework 19"
output: html_notebook
---

#Review of GSWR dplyr Concepts 

##Goals for the assignment:

1. Refresh my memory of common dplyr functions

2. Have fun with R! 

First, I will set up my workspace
```{r}
rm(list = ls())
library(dplyr)
library(here)
```

Now, I will suck in my data
```{r}
Leaves <- read.csv(here("Data", "Class_leaf_measurements.csv"))
```


Now I will examine the data
```{r}
str(Leaves) #can see that there are some messy column names
```

Now I will fix the messy column names
```{r}
names(Leaves) #I want to change "DBH..cm", "Leaf..", and "Petiole.length" so they don't have any special characters 

names(Leaves) <- c("Tree_num", "Species", "Leaf_num", "DBH_cm", "Petiole_length", "Total_length_cm", "Middle_lobe_length_cm", "Middle_lobe_width_cm", "Person")

names(Leaves) #that looks better! 
```


Now I want to convert the Species variable (which is currently a character variable) to a factor
```{r}
Leaves$Species <- as.factor(Leaves$Species)  #convert character column to factor
class(Leaves$Species) #check class, it's a factor now!
```

Now I will check how manny levels are in the Species variable
```{r}
levels(Leaves$Species) #there are 9! Uh oh, must be spelling/formatting errors
```

Now I will fix the formatting so R knows we only have 3 tree species. First, I will fix the Norway maples 
```{r}
fix <- which(Leaves$Species == "Norway mapel" | Leaves$Species == "Norway Maple") #tell R that "Norway mapel" and "Norway Maple" are wrong

Leaves$Species[fix]<- "Norway maple" #assign the wrong versions to "Norway maple"
unique(Leaves$Species) #can see that they are now labeled as "Norway maple" (even though there are still 9 levels showing...) 
```

And now I will fix sugar maples
```{r}
bad <- which(Leaves$Species == "Sugar maple" | Leaves$Species == "Sugar Maple") #tell R that "Sugar maple" and "Sugar Maple" are wrong

Leaves$Species[bad]<- "sugar maple" #assign the wrong versions to "sugar maple"
unique(Leaves$Species) #can see that they are now labeled as "sugar maple" 
```

Lastly, I will fix the red maples
```{r}
no <- which(Leaves$Species == "Red maple" | Leaves$Species == "Red Maple") #tell R that "Red maple" and "Red Maple" are wrong

Leaves$Species[no]<- "red maple" #assign the wrong versions to "red maple"
unique(Leaves$Species) #can see that they are now labeled as "red maple" 
```

Now I will create a new dataset called "Sugar" by selecting only the sugar maple data from the Leaves data set
```{r}
Sugar <- Leaves %>%
  filter(Species == "sugar maple")
```


Now I will subset the columns in Sugar using dplyr commands so that you leave out the Person, Leaf number and Tree number variables. Save the result as Sugar.
```{r}
Sugar <- select(Sugar, -c(Person, Leaf_num, Tree_num))
```

Now I will create a histogram of petiole lengths using the Sugar data set. 
```{r}
Petiole_lengths <- Sugar$Petiole_length
hist(Petiole_lengths)
```





