---
title: "Homework 12"
output: html_notebook
---

Set up my workspace 
```{r}
rm(list = ls()) #clear R's brain
library(here)
library(tidyverse)
```


Suck in the data
```{r}
Cover <- read.csv(here("Data", "donnerville_percent_cover.csv"), stringsAsFactors = TRUE) #tells R that character data are factors 
```


Glimpse the data to see the variable, variable types, number of observations, etc. 
```{r}
glimpse(Cover) #can see that there are 36 oberservations of 9 variables
```


Use `head` to view the first 6 rows of the data
```{r}
head(Cover) #everything seems in order
```


See the structure of the data
```{r}
str(Cover)
```


I now want to rename some of the columns so they follow our rules for data formatting
```{r}
Cover <- Cover %>% 
  rename(
    Plot_type = Plot.type,
    Plot_dim = Plot.dim
    )
names(Cover)
```


Now I want to determine the average percent cover by cover type (bare, bryophytes, vascular plants) for Donnerville 1
```{r}
Cover %>%
  filter(Site == "Donnerville 1") %>%
  group_by(Type) %>%
  summarize(percent_cover = mean(Percent)) 
```


Now I want to determine the average percent vascular plant cover across all sites.
```{r}
Cover %>%
  filter(Type == "Vascular") %>%
  summarize(avg_percent_vasc = mean(Percent)) #expecting to see a 1x1 data frame
```

New trick! Can use `rm(Cover)` to remove a specific data frame from the environment 

