---
title: "Chapter 2 Part 1"
output: html_notebook
---

First I will clear R's brain
```{r}
rm(list = ls())
```


Now I will run packages 
```{r}
library(dplyr)
```


Now I will import my data using read_csv() method
```{r}
library(readr)
compensation <- read_csv("Data/compensation.csv")
View(compensation)
```


Another method to import data is using the "file.choose()" function and then "reading" the data into R. 
```{r}
DF <- read.csv("/home/ohbern20/BioStats_Fall22/Bernier_Homework/Data/compensation.csv")
```


Now check that my data are my data

View names assigned to each column
```{r}
names(DF)
```

View first six rows of dataset
```{r}
head(DF)
```

See the dimensions of the dataset
```{r}
dim(DF)
```


See the structure of the dataset (what types of variables, number of obs, etc.)
```{r}
str(DF)
```

Get a horizontal view of the data
```{r}
glimpse(DF) #can see that the data have 3 variables, Root, Fruit, and Grazing
```

Get a vertical view of the data
```{r}
tbl_df(DF) #can see the data displayed in a nice vertical table 
```



