---
title: "Homework 10"
output: html_notebook
---


In this homework I will pull in and examine the squirrel.csv data using the methods we learned in chapter two of GSWR and in class

First set things up
```{r}
rm(list = ls()) #clear R's brain
library(here) #turn on libraries
library(ggplot2)
library(dplyr)
```


Now pull in the data
```{r}
DF <- read.csv(here("Data", "squirrels.csv"), stringsAsFactors = TRUE)
```


Now I will examine the data. First I will check to make sure all the variable names look correct
```{r}
names(DF) #they do! 
```

Next I will check to see what type of data R thinks each variable is, and see someof the entered data for each variable
```{r}
glimpse(DF) #I see factor (character) data and integer (numerical) data 
```


I will use the `str()` function to display the total number of observations, number of variables, and the type of data and number of levels for each variable. 
```{r}
str(DF) #this seems correct! 
```


Finally, I will display the first 6 rows of the dataset in a pretty table
```{r}
head(DF) 
```



