---
title: "Applying more dplyr to my data"
output: html_notebook
---

In this assignment, my goal is to tidy the first day of class data, inclduing changing column titles, and practice dplyr functions.

First, I will set up my workspace
```{r}
rm(list = ls())
library(tidyverse)
library(here)
```

Now I will suck in my data
```{r}
DF <- read.csv(here("Data", "day_one_data_for_homework13.csv"))
```


Now I will view my data in a new tab
```{r}
View(DF) #can see that R changed the column headers that had special characters/spaces to include a period
```


Now I will use the names() function (hint - chapter 2) to change the column names to, in order from left to right: Name, Date, Part, Tree_num, Species, leaf_length, leaf_width, leaf_units
```{r}
names(DF) #let's see the names we have

names(DF) <- c("Name", "Date", "Part", "Tree_num", "Species", "leaf_length", "leaf_width",  "leaf_units")

names(DF) #can see the all the names were changed to how I want them! 
```


Notice that some of the leaf measurements are in inches, and some are in metric units. Write code that selects just the rows of data that are in English units (inches) and that saves them to a new data frame called "English"
```{r}
English <- DF %>%
  filter(leaf_units == "inches")
```

Now notice that the dates have been entered in several formats, and that for one person, the years associated with the dates are incorrect. Write code that collects all of the data for which the dates have been formatted in the same way together into a single data frame and call it "Uniform_dates"
```{r}
Uniform_dates <- DF %>%
  filter(Date == "8/24/2022")
```



