---
title: "Using dplyr to fix mistakes"
output: html_notebook
---

In this homework I will practice dplyr skills to fix mistakes in data frames.

First, I will set up my workspace
```{r}
rm(list = ls())
library(here)
library(dplyr)
```


Now I will import my data
```{r}
DF <- read.csv(here("Data", "drug_resistance.csv")) #run without stringsAsFactors 
summary(DF) #get summary statistics for the data
```
Now I will try importing the data with `stringsAsFactors = TRUE`
```{r}
DF <- read.csv(here("Data", "drug_resistance.csv"), stringsAsFactors = TRUE)
summary(DF)
```
I can see from that when the data was imported with `stringsAsFactors = TRUE` the summary statistics told me the different levels for the variables that were initially interpreted as character/string data (Treatment and tech.name). However, when these variables were sucked in as factor data, the different spellings of each observation were interpreted as different levels of the variable.

Now I will examine the different levels of the Treatment and tech.name variables 
```{r}
levels(DF$Treatment) #I can see that there are multiple versions of each category, so I need to fix that

levels(DF$tech.name) #there are also multiple spellings here
```

Now I will clean up the Treatment variable
```{r}
fix <- which (DF$Treatment == "cantrol" | DF$Treatment == "Control") #tell R that "cantrol" and "Control" are wrong

DF$Treatment[fix]<- "control" #assign the wrong versions to "control"
unique(DF$Treatment) #can see that they are now labeled as "control" 


#now I will fix the drug2 observations
wrong <- which(DF$Treatment == "drug x 2" | DF$Treatment == "Drugx2") #tell R that "drug x 2" and "Drug2" are wrong

DF$Treatment[wrong]<- "drug2" #assign the wrong versions to "drug2"
unique(DF$Treatment) #can see that they are now labeled as "drug2" 

#and now fix drug3
bad <- which(DF$Treatment == "drug x 3" | DF$Treatment == "Drug x 3")
DF$Treatment[bad] <- "drug3"
unique(DF$Treatment) #now the only versions of treatment are control, drug2, and drug3 
```

Now I need to look at the tech.name variables
```{r}
#first let me look at the version of tech.name again, and while I'm at it let me change the name of the variable as well (and the name of colony.count)
DF <- DF %>% 
  rename(
    tech_name = tech.name,
    colony_count = colony.count)
names(DF)

levels(DF$tech_name)
```

Now I will modify the versions of tech_name
```{r}
no <- which(DF$tech_name == "becca")
DF$tech_name[no] <- "Becca"
unique(DF$tech_name)

stop <- which(DF$tech_name == "brain" | DF$tech_name == "brian")
DF$tech_name[stop] <- "Brian"
unique(DF$tech_name) #now I only see one version of each name
```


There should now only be one choice for each level of the Treatment and tech_name variables in the dataframe

