---
title: "GSWR Chapter 3 Exercise 1"
output: html_notebook
---

In this .Rmd, I will be working through the material covered in chapter 3 of Getting Started with R. I will demonstrate skills relating to data management, learn how to manipulate data, and explore different ways to familiarize myself with my data. 

First, I will set things up. 
```{r}
rm(list = ls())
library(here)
library(ggplot2)
library(dplyr)
```


Now I will suck in my data.
```{r}
compensation <- read.csv(here("Data", "compensation.csv"))
```


Now I will get a summary of the compensation data using the `summary()` function. 
```{r}
summary(compensation)
```

Now I will use `select()` to grab a specific column from the dataset
```{r}
select(compensation, Fruit) #select the Fruit column 
```

Can also grab all the columns minus specified ones
```{r}
select(compensation, -Root) #grab all columns except Root
```


What if we want to grab rows? Use `slice()`!
```{r}
slice(compensation, 2) #can see 1 row returned 
```

We could also grab rows 2-10
```{r}
slice(compensation, 2:10)
```

Can also collect specific rows using `c()` 
```{r}
slice(compensation, c(2, 3, 10))
```


Now use `with()` to tell R to look at a data frame and do a specific task
```{r}
with(compensation, Fruit > 80)
```



And now my favorite, `filter()`! I will find teh rows where it is true that Fruit is > 80 and return them as a data frame
```{r}
filter(compensation, Fruit > 80) #cool!
```


Can filter for multiple conditions
```{r}
filter(compensation, Fruit > 80 | Fruit < 20) #the "|" symbol means "OR" 
```


Now I will assign the results of the dataset filtered above to an object
```{r}
lo_hi_fruit <- filter(compensation, Fruit > 80 | Fruit < 20)
lo_hi_fruit #now look at it
```


Now before I use `mutate()` I will use `head()` to limit the rows of data
```{r}
head(compensation)
```


Now I will take the value of log(Fruit) and add it to a column called "logFruit" which will be added to the compensation dataset
```{r}
compensation <- mutate(compensation, logFruit = log(Fruit))
head(compensation) #what does compensation look like now? There's a new column called logFruit!
```


Now I will use `arrange()` to order my data in a specific way
```{r}
arrange(compensation, Fruit) #telling R to arrange the data in order of increasing value of Fruit 
```

Now I will combine multiple dplyr funcitons!
```{r}
select(filter(compensation, Fruit > 80), Root) #R will filter the dataset so that only observations where Fruit is greater than 20 exist, and tehn select the Root column for those observations
```


Now I will tell R to do the same thing as above but using the piping method
```{r}
compensation %>%
  filter(Fruit > 80) %>%
  select(Root) #same result!
```


Now to calculate some summary statistics...I will use `group_by()` and `summarise()`
```{r}
summarise(
  group_by(compensation, Grazing), 
  meanFruit = mean(Fruit))

#now assign it to a dataframe 
mean.fruit <- summarise(
  group_by(compensation, Grazing),
  meanFruit = mean(Fruit))
```

Now I will do the same task without nesting, but using piping instead
```{r}
compensation %>%
  group_by(Grazing) %>%
  summarise(meanFruit = mean(Fruit))
```


Now let's ask for a different summary statistic
```{r}
compensation %>%
  group_by(Grazing) %>%
  summarise(
    meanFruit = mean(Fruit), 
    sdFruit = sd(Fruit))
```





