---
title: "GSWR Ch. 5 Part 1"
output: html_notebook
---

First I will set up my workspace
```{r}
rm(list = ls())
library(here)
library(ggplot2)
library(tidyverse)
```


Now I will import both datasets I need to work through the text in chapter 5
```{r}
ozone <- read.csv(here("Data", "ozone.csv"))
lady <- read.csv(here("Data", "ladybirds.csv"))
```


First, I will work with the lady dataset. I will check out the data
```{r}
glimpse(lady) #I notice that I have a varialbe called "colour" that is listed in the book as "morph_colour" - just want to keep this in mind for later analysis
```


Now I want to organize the data for platting. The null hypothesis is that there is no association between the color morphs and habitat 
```{r}
totals <- lady %>% 
  group_by(Habitat, colour) %>%
  summarise(total.number = sum(number))
```


Now I will plot a simple bar graph for visualizing the data
```{r}
ggplot(totals, aes(x = Habitat, y = total.number,
                   fill = colour)) +
  geom_bar(stat = 'identity', position = 'dodge')
```


Now I want to adjust the graph colors so that they correspond to the ladybird colors!
```{r}
ggplot(totals, aes(x = Habitat, y = total.number,
                   fill = colour)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  scale_fill_manual(values = c(black = "black", red = "red")) #it looks like the black morph is more common in the industrial habitat compared to the red morph, so I am expecting to reject the null hypothesis
```

Now, before I can run the statistical model, I need to fit the data to a matix
```{r}
lady.mat <- xtabs(number ~ Habitat + colour,
                  data = lady)

lady.mat
```

Now the chi square test!
```{r}
lady.chi <- chisq.test(lady.mat) #small p-value!

names(lady.chi)

lady.chi$expected
```


      
Now, I will move on to a two-sampel t-test

First, glimpse the data
```{r}
glimpse(ozone)
```


Now, plot the data. Null hypothesis is that there is no difference in ozone levels between the two locaitons
```{r}
ggplot(ozone, aes(x = Ozone)) +
  geom_histogram(binwidth = 10) +
  facet_wrap(~Garden.location, ncol = 1) +
  theme_bw() #seems like we might reject the null hypothesis based on what this figure shows
```


Time for the t-test!
```{r}
t.test(Ozone ~ Garden.location, data = ozone) #small p-value! The level of ozone is statistically different between the two locations 
```


A slightly different version...
```{r}
var.test(Ozone ~ Garden.location, data = ozone)
```







