---
title: "Homework 26"
output: html_notebook
---

Set things up 
```{r}
rm(list = ls())
library(here)
library(tidyverse)
library(ggfortify)
```

Now let's pull in the data
```{r}
soil <- read.csv(here("Data", "soil_reduced-1.csv"), stringsAsFactors = TRUE)
```

Before we begin any analysis, we need to check our data for possible errors/outliers. Let's being with that

A. Forest names
```{r}
levels(soil$Forest) #5 levels, looks good! 
```


Let's also run a summary on all the variables 
```{r}
summary(soil) #pay attention to where max is wayyy higher than mean
```


## We'd like to ask the following questions from these data:

### 1. Is there a relationship between Soil organic matter (SOM, %) and soil calcium (Ca_ppm)?

B. Let's check soil organic matter and soil calcium

Start with a histogram of soil organic matter
```{r}
ggplot(soil, aes(x = SOM_percent)) +
  geom_histogram() #not one lone outlier, seems ok
```

Now check soil calcium
```{r}
ggplot(soil, aes(x = Ca_ppm)) +
  geom_histogram() #looks reasonable
```


Both variables are continuous 
#### Plot relationships 
```{r}
ggplot(soil, aes(x = SOM_percent, y = Ca_ppm, color = Forest)) +
  geom_point() +
  theme_bw()
```
Guess relationship: From looking at the data it appears there may be a weak positive relationship between SOM and Ca

Can I guess the slope or y intercept? 
Total rise: 1000
Total run: 53 
Slope: 1000/53 = 18ish  
Y intercept: 0ish, -5?  

What statistical model fits these data?
Both response and predictor are continuous, so linear regression seems appropriate

So, let's construct the model
```{r}
SOM_Ca_mod <- lm(Ca_ppm ~ SOM_percent, soil)
```

Now I need to check the model assumptions before I can interpret the model 
```{r}
autoplot(SOM_Ca_mod) 

#normal Q-Q plot shows that most of the data points fit normal dist
#Residuals vs fitted shows that variance of residuals is kinda unequal (drop in line at low values, rather than straight line)
```

Based on my observations of the autoplot results, I am modestly comfortable that my data meet the lm() assumptions

Now look at model outcome
```{r}
anova(SOM_Ca_mod) #asking for standard output of an anova, not calculating an anova  
```

The P-value is 0.117 which is large - accept null hypothesis (that there is no meaningful statistical relationship b/w SOM and Ca)

Now run the summary
```{r}
summary(SOM_Ca_mod) #estimates intercept as 307 and slope as 5 - different from what we guessed 
#R squared = 0.05 - only 5% of variation in calcium can be explained by abundance of SOM (even if it was significant relationship)
```

Lastly, we would plot the relationship if there were a significant relationship between the variables


### 2. Does the abundance of worms differ between Donnerville and the Kip Tract?

The null hypothesis for this question is there is no significant difference in the abundance of worms between the Donnerville and Kip Tract forests. 

I will look at the forest names first
```{r}
levels(soil$Forest) #looks good
```

Since we only want data from the Kip and Donnerville plots, I will filter the data to get only worms from the kip and donnerville forests 
```{r}
worms <- soil %>% 
  filter(Forest == "DONNERVILLE" | Forest == "KIP TRACT")

worms$Forest <- factor(worms$Forest) #will do this so R relevels the Forest variable
```

Now let's relevel
```{r}
levels(worms$Forest) #only Kip and Donnerville, good
```


Now I will get a quick visual of the variables (we already examined the summary() table of the data above)
I will create a histogram of worms first to look for outliers
```{r}
ggplot(worms, aes(x = Worms)) +
  geom_histogram()
  #nothing looks suspicious  
```


Now I will manipulate the data for plotting 
```{r}
total_worms <- worms %>% group_by(Forest) %>% summarise(total_worms = sum(Worms))
```

Now I will plot both variables together and visualize the relationship
```{r}
ggplot(total_worms, aes(x = Forest, y = total_worms)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  theme_bw() 
#identity says just plot the means
#dodge says plot them next to each other
```
Guess relationship: From looking at this graph I am guessing that there will be a significant difference between the amount of worms found in Kip vs. Donnerville. It appears there are more worms in kip. 

Now I will think about what test to run. Given the types of variables, I will run a chi square test. I will now fit the data to a matrix first
```{r}
worms.mat <- xtabs(Worms ~ Forest,
                  data = worms)

worms.mat #gives me a table
```

This is where I would check assumptions. However, count data typically fit a chi squared distribution so no assumption check necessary for our purposes. 


Now I will run a chi squared test because I am dealing with one categorical variable (with two levels) and one descrete variable 
```{r}
worms.chi <- chisq.test(worms.mat) #small p-value!
```

Now I will look at the model outcome
```{r}
worms.chi #I can see that the p value is small, so I will reject the null hypothesis
```
Interpretation: 
I reject the null hypothesis. It looks like there is a significant difference between the abundance of worms in the Kip Tract compared to Donnerville. There is an association between the number of worms and which forest they are in. 


## Next question!
### 3. Does the abundance of soil Magnesium (Mg_ppm) differ between Peavine and South Hammond?

First, I will filter the data to contain data just for Peavince and South Hammond
```{r}
mg <- soil %>% filter(Forest == "PEAVINE" | Forest == "SOUTH HAMMOND")

mg$Forest <- factor(mg$Forest) #relevel
```

Now I will plot the Mg_ppm to look for errors
```{r}
ggplot(mg, aes(x = Mg_ppm)) +
  geom_histogram() #this looks fairly reasonable, could be an outlier but nothing too extreme 
```

I will examine the data a little further
```{r}
mg$Mg_ppm #the highest value doesn't actually look suspicious given the range of data
```


Since we have already observed the Forest variable, I will move ahead with plotting the variables together and observing the relationship
```{r}
ggplot(mg, aes(x = Forest, y = Mg_ppm)) +
  geom_boxplot()
```

Guess relationship: From observing this graph, it looks like there could be a slight difference between the amount of Mg found in these forests, however there is also overlap in the amounts so it may not be significant. 

Now, I will use a use a two-sample t-test since I am dealing with one categorical variable (with two groups) and one continuous variable. 

I'm not sure how to check the assumptions for this type of test. I will plan to ask about this in class and then revise if necessary - could run this test at a lm() and check assumptions that way

I will now run the test
```{r}
t.test(Mg_ppm ~ Forest, data = mg) #large p-value! The level of Mg in the soil is not significantly different between the two locations. 
```

I accept the null hypothesis that there is no significant difference in Mg between the two forest locations. 

And for a better figure that we learned about in class
```{r}
ggplot(mg, aes(Forest, Mg_ppm)) +
  geom_boxplot() +
 geom_point(position = "jitter", size = 2, alpha = 0.5)+
     labs(y = "Magnesium, ppm")+
     theme_bw()
```


## Next question!
### 4. Does the abundance of zinc (Zn_ppm) differ among forests?

First, I will visualize the Zn_ppm variable
```{r}
ggplot(soil, aes(x = Zn_ppm)) +
  geom_histogram() #there is a wide range in the data but nothing that makes me uncomfortable given that there is not one lone outlier
```

Next I will construct a visual representation of both variables to examine the relationship
```{r}
ggplot(soil, aes(x = Forest, y = Zn_ppm)) +
  geom_boxplot() #looks like there are some differences, and could be an outlier in degrasse because the highest value is way larger than the mean
```

Guess relationship: Prediction is that there is a significant difference in the levels of Zn between the forests.

Now I will fit the model for an ANOVA since I am working with a continuous variable and a categorical variable with more than two groups. 
```{r}
model_zn <- lm(Zn_ppm ~ Forest, data = soil)
```


And now I will check the assumptions
```{r}
autoplot(model_zn)  
#normal Q-Q plotL shows that most of the data points fit normal dist, with some larger values falling away from the pattern
#Residuals vs fitted: shows that variance of residuals is fairly equal (slight drop off in line at higher values, rather than straight line)
```

Given my purpose, I feel fairly confident moving on in the statistical test

I will now run the one-way ANOVA 
```{r}
anova(model_zn) #large p-value, so I accept my null hypothesis
```


I will also run the summary table for more statistics
```{r}
summary(model_zn)
```

The large P-value leads me to accept my null hypothesis. There is no statistical difference in the levels of Zn between the different forests. 

THIS RMD FILE HAS THE LEVEL OF DETAIL ON INTERPRETATION THAT I WAS LOOKING FOR IN HOMEWORK 25. 
