---
title: "GSWR Ch. 5 Part 2"
output: html_notebook
---

First I'll set up my workspace
```{r}
rm(list = ls())
library(here)
library(ggplot2)
library(tidyverse)
```

Now I will suck in my data
```{r}
plant_gr <- read.csv(here("Data", "plant.growth.rate.csv"))
```

Now I will glimpse the data
```{r}
glimpse(plant_gr)
```


Now I will get a quick visual of the data
```{r}
ggplot(plant_gr,
       aes(x = soil.moisture.content, y = plant.growth.rate)) +
  geom_point() +
  ylab("Plant Growth Rate (mm/week)") +
  theme_bw()

#notice there is a positive slop to the data
```

Now I will fit the model for a linear regression where plant growth rate is a function of soil moisture content
```{r}
model_pgr <- lm(plant.growth.rate ~ soil.moisture.content, 
                data = plant_gr)
```

Now I will check the assumptions. First I need to install ggfortify
```{r}
library(ggfortify)

#now I will use autoplot to look at assumptions
autoplot(model_pgr, smooth.colour = NA)
```

And now for the interpretation
```{r}
anova(model_pgr) #can see the large F-value and small degrees freedom 
```

And now run the summary table
```{r}
summary(model_pgr)
```

And now to go back to a figure that shows the relationship between the variables
```{r}
ggplot(plant_gr, aes(x = soil.moisture.content, y = plant.growth.rate)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  ylab("Plant Growth Rate (mm/week)") +
  theme_bw()
  #makes a nice linear model fitted line on top of the data points 
```

Now for a one-way ANOVA 
First, upload the data I need 
```{r}
daphnia <- read.csv(here("Data", "Daphniagrowth.csv"))
```

And glimpse the data
```{r}
glimpse(daphnia) #predict that parasite presnse will negatively impact growth rate
```

And now build a figure
```{r}
ggplot(daphnia, aes(x = parasite, y = growth.rate)) +
  geom_boxplot() +
  theme_bw() +
  coord_flip() #and flip coordinates so easier to read
#looks like the prediciton is on the right track
```

Construct the ANOVA next
```{r}
model_grow <- lm(growth.rate ~ parasite, data = daphnia)
```

And check the assumptions
```{r}
autoplot(model_grow, smooth.colour = NA)
```

And not make an inference from a one-way anova
```{r}
anova(model_grow) #large F-value, thus small p value and can reject null hypothesis (there is no difference between growth rates)
```

Summary table
```{r}
summary(model_grow)
```

And now, calculate the actual mean growth rate for the different treatments
```{r}
sumData <- daphnia %>%
  group_by(parasite) %>%
  summarise(meanGR = mean(growth.rate))
```




