---
title: "More with GGPlot"
output: html_notebook
---

First, I'll set up my workspace
```{r}
rm(list = ls())
library(ggplot2)
```

## Chapter 1 code
```{r}
#install.packages(c(
 #"ggnewscale", "ggplot2", "ggraph", "ggrepel", "ggtext", "ggthemes", 
  #"hexbin", "Hmisc", "mapproj", "maps", "munsell", "ozmaps", 
  #"paletteer", "patchwork", "rmapshaper", "scico", "seriation", "sf", 
  #"stars", "tidygraph", "tidyr", "wesanderson" )) 

#that almost broke my computer, I would rather run these separately as I need them so I will make this code a comment instead 
```

## Chapter 2 code

Access some built-in data by loading ggplot
```{r}
library(ggplot2)

mpg
```

## Scatter plots first

Make a simple scatter plot graph 
```{r}
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point()

#data = mpg, aesthetica = engine size to x, fuel economy to y, layer = points
```

Alternatively, it is assumed that the first two objects in aes() are mapped to x and y, so it doesn't need to be stated
```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point()
```

What does this graph show? Is it useful? 
```{r}
ggplot(mpg, aes(model, manufacturer)) + geom_point() #not very useful because can't read the x axis, but it's trying to show which type of car is made by which manufacturers 
```


My prediction is that this will show the trend of fuel consumption in the city vs the country 
```{r}
ggplot(mpg, aes(cty, hwy)) + geom_point()
```

Prediction: as carat increases, price will increase
```{r}
ggplot(diamonds, aes(carat, price)) + geom_point()
```

Prediction: date and unemployment will fluctuate 
```{r}
ggplot(economics, aes(date, unemploy)) + geom_line()
```

Prediction: distribution of mpg in the city, most cars will get lower mileage than higher mieage
```{r}
ggplot(mpg, aes(cty)) + geom_histogram()
```

Now make it pretty: map mpg for highway by displacement with car types color coded 
```{r}
ggplot(mpg, aes(displ, hwy, colour = class)) + 
  geom_point()
```

Compare:
```{r}
ggplot(mpg, aes(displ, hwy)) + geom_point(aes(colour = "blue")) #this graph is done by color, but it is just named "blue," not acutally telling it to plot the color blue


ggplot(mpg, aes(displ, hwy)) + geom_point(colour = "blue") #this graph the points are actually blue, though not labelled blue
```

Now, faceting
```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  facet_wrap(~class) #make a graph for each of the levels of "class"
```

Add a smoother to a plot
```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  geom_smooth()
#> `geom_smooth()` using method = 'loess' and formula 'y ~ x

#Adds a line to fit the points 
```

More specifics...wiggliness of line controlled by span
```{r}
ggplot(mpg, aes (displ, hwy)) + 
  geom_point() + 
  geom_smooth(span = 0.2)
#very wiggly
```


```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  geom_smooth(span = 1) #less wiggly
```


Now I need to run a new package
```{r}
install.packages("nlme")
library(mgcv)
```

And now a new graph using mgcv package to fit a generalized model additive
```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  geom_smooth(method = "gam", formula = y ~ s(x))
```

Use lm() to fit a linear model and give the line the best fit
```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```


## Now onto boxplots

A simple one
```{r}
ggplot(mpg, aes(drv, hwy)) + 
  geom_point()
```

Some cool ones
```{r}
ggplot(mpg, aes(drv, hwy)) + geom_jitter()

ggplot(mpg, aes(drv, hwy)) + geom_boxplot()

ggplot(mpg, aes(drv, hwy)) + geom_violin()
```

## Now some histograms/frequency plots
```{r}
ggplot(mpg, aes(hwy)) + geom_histogram()
#> `stat_bin()` using `bins = 30`. Pick better value with `binwidth`.

ggplot(mpg, aes(hwy)) + geom_freqpoly()
#> `stat_bin()` using `bins = 30`. Pick better value with `binwidth`.
```


Play around with binwidth
```{r}
ggplot(mpg, aes(hwy)) + 
  geom_freqpoly(binwidth = 2.5)

ggplot(mpg, aes(hwy)) + 
  geom_freqpoly(binwidth = 1)
```


```{r}
ggplot(mpg, aes(displ, colour = drv)) +  #all info on one graph, by color
  geom_freqpoly(binwidth = 0.5)

ggplot(mpg, aes(displ, fill = drv)) + #separate into three graphs
  geom_histogram(binwidth = 0.5) + 
  facet_wrap(~drv, ncol = 1)
```


## Bar charts
```{r}
ggplot(mpg, aes(manufacturer)) + 
  geom_bar() #I can't read the x axis!
```


Build some data...
```{r}
drugs <- data.frame(
  drug = c("a", "b", "c"),
  effect = c(4.2, 9.7, 6.1)
)
```


And plot it with two methods
```{r}
ggplot(drugs, aes(drug, effect)) + geom_bar(stat = "identity")

ggplot(drugs, aes(drug, effect)) + geom_point()
```

Similar graphs convey different information
```{r}
ggplot(economics, aes(date, unemploy / pop)) +
  geom_line() #shows unemployment rate

ggplot(economics, aes(date, uempmed)) +
  geom_line() #shows median number of weeks unemployed
```


Now plot unemployment rate by length of unemployment
```{r}
ggplot(economics, aes(unemploy / pop, uempmed)) + 
  geom_path() +
  geom_point() #basic version

year <- function(x) as.POSIXlt(x)$year + 1900

ggplot(economics, aes(unemploy / pop, uempmed)) + 
  geom_path(colour = "grey50") +
  geom_point(aes(colour = year(date))) #add color gradient by time (years)
```

Modifying axes
```{r}
ggplot(mpg, aes(cty, hwy)) +
  geom_point(alpha = 1 / 3) #play with alpha

ggplot(mpg, aes(cty, hwy)) + #add axis labels 
  geom_point(alpha = 1 / 3) + 
  xlab("city driving (mpg)") + 
  ylab("highway driving (mpg)")

# Remove the axis labels with NULL
ggplot(mpg, aes(cty, hwy)) +
  geom_point(alpha = 1 / 3) + 
  xlab(NULL) + 
  ylab(NULL)
```

Modify limits of the plots
```{r}
ggplot(mpg, aes(drv, hwy)) +
  geom_jitter(width = 0.25)

ggplot(mpg, aes(drv, hwy)) +
  geom_jitter(width = 0.25) + 
  xlim("f", "r") + 
  ylim(20, 30)
#> Warning: Removed 140 rows containing missing values (geom_point).
  
# For continuous scales, use NA to set only one limit
ggplot(mpg, aes(drv, hwy)) +
  geom_jitter(width = 0.25, na.rm = TRUE) + 
  ylim(NA, 30)
```


Save a plot
```{r}
p <- ggplot(mpg, aes(displ, hwy, colour = factor(cyl))) +
  geom_point()

#and view it 
print(p)
```

Use ggsave method
```{r}
# Save png to disk
ggsave("plot.png", p, width = 5, height = 5)
```


Use summary() to describe structure of plot
```{r}
summary(p)
```


Save a copy of the plot so you can recreate it later
```{r}
saveRDS(p, "plot.rds")
q <- readRDS("plot.rds")
```



