---
title: "GSWR Chapter 4"
output: html_notebook
---

First, set up my workspace
```{r}
rm(list = ls())
library(ggplot2)
library(tidyverse)
library(here)
```

Now suck in the compensation data
```{r}
compensation <- read.csv(here("Data", "compensation.csv"))
```

Now check out the data
```{r}
glimpse(compensation)
```

Now I will make my first ggplot picture - a scatter plot
```{r}
ggplot(compensation, aes(x = Root, y = Fruit)) +
  geom_point()
```

Now I want to add a color theme
```{r}
ggplot(compensation, aes(x = Root, y = Fruit)) +
  geom_point() +
  theme_bw()
```

Now I want to increase the size of the points, add axis labels, and make it so that the colors of the points match the levels in the group
```{r}
ggplot(compensation, aes(x = Root, y = Fruit, color = Grazing)) + #color = to specify color matches levels
  geom_point(size = 5) + #specify point size
  xlab("Root Biomass") + #add x label
  ylab("Fruit Production") + #add y label
  theme_bw()
```

Could also change the shape of the points to correspond to the grazing level
```{r}
ggplot(compensation, aes(x = Root, y = Fruit, shape = Grazing)) +
  geom_point(size = 5) + 
  xlab("Root Biomass") + 
  ylab("Fruit Production") + 
  theme_bw()
```



Now, moving on to box and whisker plots...
```{r}
ggplot(compensation, aes(x = Grazing, y = Fruit)) +
  geom_boxplot() +
  xlab("Grazing Treatment") +
  ylab("Fruit Production") +
  theme_bw()
```


Now, I'll add some raw data points over the boxplot
```{r}
ggplot(compensation, aes(x = Grazing, y = Fruit)) +
  geom_boxplot() +
  geom_point(size = 4, color = 'lightgrey', alpha = 0.5) + #add points with specific size, color, and transparency 
  xlab("Grazing Treatment") +
  ylab("Fruit Production") +
  theme_bw()
```


Now lets make a histogram!
```{r}
ggplot(compensation, aes(x = Fruit)) +
  geom_histogram()
```


Now let's make it look a little nicer
```{r}
ggplot(compensation, aes(x = Fruit)) +
  geom_histogram(bins = 10) #use 10 bars
ggplot(compensation, aes(x = Fruit)) +
  geom_histogram(binwidth = 15) #adjust the width of each bar
```


Also...we can facet
```{r}
ggplot(compensation, aes(x = Fruit)) +
  geom_histogram(binwidth = 15) +
  facet_wrap(~Grazing) #tells R to make a histrogram for each level of Grazing 
```

Lastly, if I wanted to save these plots anywhere as a png file, I could use ggsave()







