---
title: "More with GGPlot"
output: html_notebook
---

First, I'll set up my workspace
```{r}
rm(list = ls())
```

## Chapter 1 code
```{r}
install.packages(c(
  "colorBlindness", "directlabels", "dplyr", "ggforce", "gghighlight", 
  "ggnewscale", "ggplot2", "ggraph", "ggrepel", "ggtext", "ggthemes", 
  "hexbin", "Hmisc", "mapproj", "maps", "munsell", "ozmaps", 
  "paletteer", "patchwork", "rmapshaper", "scico", "seriation", "sf", 
  "stars", "tidygraph", "tidyr", "wesanderson" 
)) #that almost broke my computer
```

## Chapter 2 code

Access some built-in data by loading ggplot
```{r}
library(ggplot2)

mpg
```


Make a simple scatter plot graph 
```{r}
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point()

#data = mpg, aesthetica = engine size to x, fuel economy to y, layer = points
```

Alternatively, it is assumed that the first two objects in aes() are mapped to x and y, so it doesn't need to be stated
```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point()
```

What does this graph show? Is it useful? 
```{r}
ggplot(mpg, aes(model, manufacturer)) + geom_point() #not very useful because can't read the x axis, but it's trying to show which type of car is made by which manufacturers 
```


My prediction is that this will show the trend of fuel consumption in the city vs the country 
```{r}
ggplot(mpg, aes(cty, hwy)) + geom_point()
```

Prediction: as carat increases, price will increase
```{r}
ggplot(diamonds, aes(carat, price)) + geom_point()
```

Prediction: date and unemployment will fluctuate 
```{r}
ggplot(economics, aes(date, unemploy)) + geom_line()
```

Prediction: distribution of mpg in the city, most cars will get lower mileage than higher mieage
```{r}
ggplot(mpg, aes(cty)) + geom_histogram()
```

Now make it pretty: map mpg for highway by displacement with car types color coded 
```{r}
ggplot(mpg, aes(displ, hwy, colour = class)) + 
  geom_point()
```

Compare:
```{r}
ggplot(mpg, aes(displ, hwy)) + geom_point(aes(colour = "blue")) #this graph is done by color, but it is just named "blue," not acutally telling it to plot the color blue


ggplot(mpg, aes(displ, hwy)) + geom_point(colour = "blue") #this graph the points are actually blue, though not labelled blue
```

Now, faceting
```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  facet_wrap(~class) #make a graph for each of the levels of "class"
```

Add a smoother to a plot
```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  geom_smooth()
#> `geom_smooth()` using method = 'loess' and formula 'y ~ x

#Adds a line to fit the points 
```

More specifics...
```{r}
ggplot(mpg, aes (displ, hwy)) + 
  geom_point() + 
  geom_smooth(span = 0.2)

#specify how wide the line spans by 'span = '
```



```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  geom_smooth(span = 1)
```








